Certified Symbolic Finite Transducers: Formalization and Applications to String Analysis

### 2025-09-10 Paper/introduction.tex
Typos, more discussion added

Despite these, existing verified implementations for string analysis do not
handle the following indispensable string operations for analysis of web
applications (e.g. see
\cite{DBLP:conf/popl/LinB16,Kern,Berkeley-JavaScript,DBLP:conf/sp/SaxenaAHMMS10,uss/HooimeijerLMSV11}):
\emph{string transductions} and \emph{replaceall} operators. In particular, 
string transductions are used frequently in web applications (e.g. for
sanitization, as well as implicitly applied transductions, for instance
through \texttt{innerHTML}) and are well-known to be critical for analyzing XSS
vulnerabilities for web applications
\cite{Berkeley-JavaScript,DBLP:conf/sp/SaxenaAHMMS10,Kern,uss/HooimeijerLMSV11,DBLP:conf/popl/LinB16}. In particular, SMT-LIB string constraints that 
arise from such applications require the \emph{replaceall} operator (part of
SMT-LIB), which is currently not handled by any existing verified
implementation. This paper tackles the formalization of SFA/SFT and
their associated algorithms, as well as applications to analysis of sanitizers
and verified solver for string constraints. 
%\textcolor{red}{Updated up to here
%so far}

@@

To evaluate the effectiveness and efficiency of our formalization, we 
provide two applications: (1) analysis of the web application sanitizers for html, css, json, and javascript
which are used for preventing XSS attacks, and (2) an SMT-LIB \cite{smtlib}
compliant certified string solver handling equality, concatenation, regular
constraints, and more importantly the string replaceall operator, which has been
argued to be indispensable for analysis of XSS 
\cite{DBLP:conf/popl/LinB16,Kern,Berkeley-JavaScript,DBLP:conf/sp/SaxenaAHMMS10,uss/HooimeijerLMSV11}.
The experimental results demonstrate that our formalization achieves computational efficiency in constraint-solving scenarios.

>>>>

Despite these, existing verified implementations for string analysis do not
handle the following indispensable string operations for analysis of web
applications (e.g. see
\cite{DBLP:conf/popl/LinB16,Kern,Berkeley-JavaScript,systematic-transduction,uss/HooimeijerLMSV11}):
\emph{string transductions} and \emph{replaceall} operators. In particular, 
string transductions are used frequently in web applications (e.g. for
sanitization, as well as implicitly applied transductions, for instance
through \texttt{innerHTML}) and are well-known to be critical for analyzing XSS
vulnerabilities for web applications
\cite{Berkeley-JavaScript,systematic-transduction,Kern,uss/HooimeijerLMSV11,DBLP:conf/popl/LinB16}. In particular, SMT-LIB string constraints that 
arise from such applications require the \emph{replaceall} operator (part of
SMT-LIB), which is currently not handled by any existing verified
implementation. This paper tackles the formalization of SFA/SFT and
their associated algorithms, as well as applications to analysis of sanitizers
and verified solver for string constraints. 
%\textcolor{red}{Updated up to here
%so far}

@@

To evaluate the effectiveness and efficiency of our formalization, we 
provide two applications: (1) analysis of the web application sanitizers for html, css, json, and javascript
which are used for preventing XSS attacks, and (2) an SMT-LIB \cite{smtlib}
compliant certified string solver handling equality, concatenation, regular
constraints, and more importantly the string replaceall operator, which has been
argued to be indispensable for analysis of XSS 
\cite{DBLP:conf/popl/LinB16,Kern,Berkeley-JavaScript,systematic-transduction,uss/HooimeijerLMSV11}.
The experimental results demonstrate that our formalization achieves computational efficiency in constraint-solving scenarios.
